---
title: "Maps, how do they work?"
subtitle: 'Source: https://github.com/ASRCsoft/map_backends'
author: "William May"
date: "`r gsub(' 0', ' ', format(Sys.Date(), '%B %d, %Y'))`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(magrittr)
library(DiagrammeR)
library(leaflet)
knitr::opts_chunk$set(echo = FALSE)
```

<style>
/* 2-column formatting */w
.columns-2 {
  height: 76%;
  -webkit-column-gap: 10%;
  -moz-column-gap: 10%;
  column-gap: 10%;
}
.forceBreak {
  -webkit-column-break-after: always;
  break-after: column;
}
/* Definition lists */
dt {
    font-weight: bold;
    margin-top: 30px;
    margin-bottom: 10px;
}
dt::after {
  content: ":";
}
dd {
    margin-left: 20px;
}
</style>

## Outline

```{r}
node_labels = c('HYSPLIT/IDEA/etc', 'Python/GDAL\n(parse raw data)', 'Postgres (database)', 'Flask (server)',
								'Leaflet (map)')
n_nodes = length(node_labels)
nodes = create_node_df(n_nodes, label = node_labels)
edges = create_edge_df(1:4, 2:5)
create_graph(nodes, edges, attr_theme = NULL) %>%
	render_graph
```

## Flask

- listens for API calls
- gets data from postgres
- sends back json and png files

## Leaflet {.columns-2}

- flexible and easily extensible
- [Leaflet.TimeDimension](https://github.com/socib/Leaflet.TimeDimension) plugin provides time controls (but sometimes it's a pain)
- can also be used via R and Python libraries

<p class="forceBreak"></p>

```{r out.width="100%"}
asrc_lat = 42.692134
asrc_lng = -73.832040
leaflet() %>%
	setView(asrc_lng, asrc_lat, zoom = 16) %>%
  addTiles() %>%
  addMarkers(lng = asrc_lng, lat = asrc_lat, popup = 'ASRC')
```

## Use Shiny instead {.columns-2}

- runs its own server (no flask required)
- easy to make user interfaces
- easy to include interactive plots (like leaflet)
- can set up time series animations

<p class="forceBreak"></p>

```{r out.width="100%"}
node_labels = c('HYSPLIT/IDEA/etc', 'Python/GDAL\n(parse raw data)', 'Postgres (database)', 'R/Shiny')
n_nodes = length(node_labels)
nodes = create_node_df(n_nodes, label = node_labels)
edges = create_edge_df(1:3, 2:4)
create_graph(nodes, edges, attr_theme = NULL) %>%
	render_graph
```

## Python + GDAL

- GDAL (Geospatial Data Abstraction Library) needed to convert satellite swaths to standard map projections
- cron runs Python scripts that convert data and insert it into postgres

## Postgres + PostGIS

### PostgreSQL (aka Postgres)

- old open source database with a lot of nice features

### PostGIS

- adds geospatial types (points, rasters, polygons, etc) to postgres
- can convert data to png, geotiff, and geojson

## human issues

Sarah Lu group/DEC does not know what they want, just ask for "demos" and new features over and over again. Try to get a list of

- what data they're interested in
- what attributes they want to see (location, height, concentration, ...)
- which datasets they want to directly compare

and be ready to make changes.

Save yourself some headaches by splitting features into multiple apps or pages instead of one ultra-complicated app.

## Inventory

HYSPLIT
  ~ Trajectory and dispersion model, uses HRRR model output (Mark Beauharnois maintains part of the code)

IDEA
  ~ Trajectory model, uses VIIRS satellite swaths and HRRR
  
CMAQ
  ~ Air quality model (no database or server yet)
  
Lidar
  ~ Leosphere Wind Lidar, instrument that measures wind speeds and aerosols

## useful links

- [etl](https://github.com/beanumber/etl) R package, convenient way to organize databases in R
- [leaflet](https://github.com/rstudio/leaflet) R package
- [leaflet.extras](https://github.com/bhaskarvk/leaflet.extras) R package
- [Leaflet.LayerArray](https://github.com/ASRCsoft/Leaflet.LayerArray): ASRC extension of Leaflet.TimeDimension
- [MobilityDB](https://github.com/ULB-CoDE-WIT/MobilityDB): a Postgres extension with trajectory functions
- [rpostgis](https://github.com/mablab/rpostgis) R package
- [Shiny](https://github.com/rstudio/shiny)
- [Spatial](https://cran.r-project.org/view=Spatial) R task view
- [Spatiotemporal](https://cran.r-project.org/view=SpatioTemporal) R task view
- [splitR](https://github.com/rich-iannone/splitr): R HYSPLIT package
- [Tracking](https://cran.r-project.org/view=Tracking) R task view
